---
title: "Practical Machine Learning Assignment"
author: "Tanishq Singla"
date: "2 September 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the final project for the "*Practical Machine Learning*" course. 

## Summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

The training data for this project is available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv).
The test data for the project is available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv).

The goal of the project is to predict the manner in which they did the exercise. This is the “classe” variable in the training set.

##Data Loading and Processing
```{r message=FALSE, warning=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(corrplot)
library(gbm)
```

##Getting the Data
```{r}
pml_training <- read.csv("./pml-training.csv", header = TRUE)
pml_testing <- read.csv("./pml-testing.csv", header = TRUE)
dim(pml_training)
```
```{r}
dim(pml_testing)
```
As shown below there are 19622 observations and 160 variables in the Training dataset.

##Cleaning the Data
Removing variables with missing values
```{r}
trainData<- pml_training[, colSums(is.na(pml_training)) == 0]
validData <- pml_testing[, colSums(is.na(pml_testing)) == 0]
dim(trainData)
```
```{r}
dim(validData)
```

###We also remove the first seven variables as they have little impact on "*classe*" variable.
```{r}
trainData <- trainData[, -c(1:7)]
validData <- validData[, -c(1:7)]
dim(trainData)
```
```{r}
dim(validData)
```

##Preparing the dataset for prediction
Splitting the training dataset into train data and test data in the ratio 7:3.
```{r}
set.seed(1234) 
inTrain <- createDataPartition(trainData$classe, p = 0.7, list = FALSE)
trainData <- trainData[inTrain, ]
testData <- trainData[-inTrain, ]
trainData$classe <- as.factor(trainData$classe)
testData$classe <- as.factor(testData$classe)
dim(trainData)
```
```{r}
dim(testData)
```

##Cleaning even further by removing the variables that are near-zero-variance
```{r}
NZV <- nearZeroVar(trainData)
trainData <- trainData[, -NZV]
testData  <- testData[, -NZV]
dim(trainData)
```
```{r}
dim(testData)
```


##Correlation
The following correlation plot uses the following parameters (source:CRAN Package ‘corrplot’) “FPC”: the first principal component order. “AOE”: the angular order tl.cex Numeric, for the size of text label (variable names) tl.col The color of text label.
```{r}
cor_mat <- cor(trainData[, -53])
corrplot(cor_mat, order = "FPC", method = "color", type = "upper", 
         tl.cex = 0.7, tl.col = rgb(0, 0, 0))
```

Finding highly correlated variables using findCorrelation function at a cutoff 0f 0.75.
```{r}
highlyCorr = findCorrelation(cor_mat, cutoff=0.75)
```

Getting the names of the variables
```{r}
names(trainData)[highlyCorr]
```

##Model
We use the Random Forests Model as it gave the best result.
```{r}
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modRF1 <- train(classe ~ ., data=trainData, method="rf", trControl=controlRF)
modRF1$finalModel
```

##Prediction
```{r}
predictRF1 <- predict(modRF1, newdata=testData)
cmrf <- confusionMatrix(predictRF1, testData$classe)
cmrf
```
###The accuracy rate using the random forest is very high: Accuracy : 1 and therefore the out-of-sample-error is equal to 0***. But it might be due to overfitting.

Lets plot the model
```{r}
plot(modRF1)
```


##Applying the model on validate data
```{r}
Results <- predict(modRF1, newdata=validData)
Results
```

